---
title: "APR p-value comparison"
author: "LeAnn Lo"
date: "January 24, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this is to compare the non corrected APR data and the batch effect removed data. The original DESeq output is sorted by p-adj values; for this comparison, the DESeq output must be sorted by p-value. This comparison will go over all of the pairwise comparisons done in APR 

# Load libraries
```{r}
suppressMessages(library(RColorBrewer))
suppressMessages(library(DESeq2))
suppressMessages(library(affy))
suppressMessages(library("vsn"))
suppressMessages(library("ggplot2"))
suppressMessages(library(ggrepel))
suppressMessages(library(genefilter))
```

## Sib vs 11C

Here we sort by p-value, a batch effect correction was made for the Sib vs 11C comparison. Let's load the resulting DESeq2 differential exression output and compare the rankings based on the p-values. These list have been sorted by raw p-values.
```{r}
# load data and only get first 7 columns, the rest are counts which we don't need

## no correction

Sib_11C_no_cor <- read.csv("/Volumes/CaChannel/RNA_SEQ/Analysis/APRotations/161221_Hisat2/Taejoon_JGlv9/Differential_Expression/Analysis_Workspace/18/uncollapsed/uncorrected_APR_Selfie/Sib_11C_stage18_DESeq2.csv")
# reorder by p-value
Sib_11C_no_cor <- Sib_11C_no_cor[order(Sib_11C_no_cor$pvalue),] 
Sib_11C_no_cor <- Sib_11C_no_cor[,1:7]

## correction
Sib_11C_cor <- read.csv("/Volumes/CaChannel/RNA_SEQ/Analysis/APRotations/161221_Hisat2/Taejoon_JGlv9/Differential_Expression/Analysis_Workspace/18/Batch_corrected/uncollapsed/Sib_11C_stage18_DESeq2.csv")
# reorder by p-value
Sib_11C_cor <- Sib_11C_cor[order(Sib_11C_cor$pvalue),] 
Sib_11C_cor <- Sib_11C_cor[,1:7]

# get length of both, they should be the same
length(Sib_11C_no_cor$Gene)
length(Sib_11C_cor$Gene)
```

Check the headers to ensure that they are sorted by raw p-value, not p-adj like the original output of DESeq.
```{r}
head(Sib_11C_no_cor)
head(Sib_11C_cor)
```

Pull out all of the genes with a p-value less than 0.05 for significantly differentially expressed genes. 
```{r}
# Significantly differentially expressed in non corrected
Sib_11C_no_cor_sig <- subset(Sib_11C_no_cor, padj < 0.05)
length(Sib_11C_no_cor_sig$Gene)

# Significantly differentially expressed in corrected
Sib_11C_cor_sig <- subset(Sib_11C_cor, padj < 0.05)
length(Sib_11C_cor_sig$Gene)
```

We can compare the ranking of the first 238 genes from the non correction list to the correction list using a spearman's correlation test. Here a test statistic of 1 is a perfect match. First, set up the data to run the correlation.
```{r}
# get first 238 genes ranked by raw p-value in non correction list
Sib_11C_no_cor_siggene <- as.character(Sib_11C_no_cor$Gene)[1:238]

# get first 238 genes ranked by raw p-value in correction list
Sib_11C_cor_siggene <- as.character(Sib_11C_cor$Gene)[1:238]

# combine these vectors into a single matrix
mat <- matrix(c(Sib_11C_no_cor_siggene, Sib_11C_cor_siggene), nrow = 238, ncol = 2)

# take a look
head(mat)
```

We can see how many genes are shared between the lists. Only 79 of these genes are shared between the two lists. 
```{r}
length(intersect(Sib_11C_no_cor_siggene, Sib_11C_cor_siggene))
```

We can test their correlation by running the Spearman correlation test. The correlation test value we receive is -0.08252794, this suggests that the lists are not closely correlated. This makes sense considering that there are only 79 genes that intersect.
```{r}
# convert matrix to numeric matrix so we can run a correlation
mat <- matrix(as.numeric(as.factor(c(Sib_11C_no_cor_siggene, Sib_11C_cor_siggene))), nrow = 238, ncol = 2)
head(mat)

# compare these ranked lists, columns 1 and 2 using spearmans corelation
cor(mat[,1], mat[,2], use="complete.obs", method = "spearman")
```

We can obtain the list of shared genes between the corrected and non corrected data. There are 79 shared genes between the two. 
```{r}
sharedgenes <- intersect(Sib_11C_no_cor_siggene, Sib_11C_cor_siggene)
sharedgenes
```













