---
title: "11C versus APR Sibs and Selfie Sibs"
author: "LeAnn Lo"
date: "February 28, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading Count Data
We are using APR sib, Selfie sib, and APR 11C data from stage 18. 
```{r countData}
suppressMessages(library(RUVSeq))
suppressMessages(library("ggplot2"))
suppressMessages(library(ggrepel))
suppressMessages(library(factoextra))
suppressMessages(library(FactoMineR))
suppressMessages(library(genefilter))

files_sib_rot <- list.files("/Volumes/cachannel/RNA_SEQ/Analysis/APRotations/Quantification/170802_HTSeq-count_Output_exon/Rotations_Batch/Stage_18/S18_Sib", recursive = TRUE, full.name=TRUE)
files_sib_selfie <- list.files("/Volumes/cachannel/RNA_SEQ/Analysis/APRotations/Quantification/170802_HTSeq-count_Output_exon/Selfies_Batch/Stage_18/selfie_S18_Sib", recursive = TRUE, full.name=TRUE)
files_apr_11C <- list.files("/Volumes/cachannel/RNA_SEQ/Analysis/APRotations/Quantification/170802_HTSeq-count_Output_exon/Rotations_Batch/Stage_18/S18_11C", recursive = TRUE, full.name=TRUE)

files <- c(files_sib_rot, files_sib_selfie, files_apr_11C)

# sample name rotations 
sampNum_rot = c("S18_Sib_1","S18_Sib_2","S18_Sib_3","S18_Sib_4","S18_Sib_5")

# sample name selfies
sampNum_self = c("S18_SibSelfie_1", "S18_SibSelfie_2")

# sample name APR 11C
sampNum_11C = c("S18_11C_1","S18_11C_2","S18_11C_3","S18_11C_4","S18_11C_5")

# combine sample names
sampNum <- c(sampNum_rot, sampNum_self, sampNum_11C)

# setting batch
batch <- as.factor(rep(c("Sibs", "11C"), c(7, 5)))

sTable = data.frame(sampleName = sampNum, fileName = files, batch = batch)
```

## DESeq #1

```{r}
suppressMessages(library(DESeq2))
design <- ~ batch
dds18 <- DESeqDataSetFromHTSeqCount(sampleTable = sTable, directory = "", design = design)
dds18 <- dds18[rowMeans(counts(dds18)) >= 5,] 
dds <- DESeq(dds18)
```

## Original Plot

```{r original plot}
suppressMessages(library(limma))

project.pca <- prcomp(t(counts(dds, normalized = TRUE)))
summary(project.pca)
project.pca.proportionvariances <- ((project.pca$sdev^2) / (sum(project.pca$sdev^2)))*100
```

```{r}
library(RColorBrewer)
colors <- brewer.pal(3, "Set2")

#Plots scatter plot for PC 1 and 2
plot(project.pca$x, type="n", main="Original PCA Plot", xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"), ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"))
points(project.pca$x, col=colors[batch], pch=19)
text(project.pca$x, labels=sampNum, cex= 0.7)
```

```{r getting avaiable comparisons from DESeq}
head(dds)
#get name of comparisons available to you
resultsNames(dds)
```

Write output function
```{r write output function}
# write output
writeOutput <- function(res, dds, name, cond, dir) {
  resOrdered <- res[order(res$padj),]
  resdata <- merge(as.data.frame(resOrdered), as.data.frame(counts(dds, normalized=TRUE)), by="row.names", sort=FALSE) #includes normalized counts in output csv
  names(resdata)[1] <- "gene" # set header of first column
  outfile <- paste(cond[1], cond[length(cond)], name, "DESeq2.csv", sep = "_")
  outfile <- paste(dir, outfile, sep = "")
  write.csv(as.data.frame(resdata), file = outfile, row.names = FALSE)
}
# setting dir
dir <- "/Volumes/cachannel/SCRIPTS/R/APRotations_RNA-Seq/Batch_Correction/11C_Sibs/"
```

# DESeq comparison of batches
There are about 1000 DE genes before any correction (SVA)
```{r}
# get results of batch vs selfie
res <- results(dds, alpha = 0.05, name = "batch_Sibs_vs_11C")
hist(res$pvalue, main = "All Sibs vs APR 11C", xlab = "p-values", col = "lavender")
summary(res)
#writeOutput(res, dds, "stage18_before_corr", c("Sibs", "APR_11C"), dir)
```



```{r SVA}
# Loading all variables 
library(sva)
dat  <- counts(dds, normalized = TRUE)
idx  <- rowMeans(dat) > 1
dat  <- dat[idx, ]
mod  <- model.matrix(~ batch, colData(dds))
mod0 <- model.matrix(~ 1, colData(dds))
svseq <- svaseq(dat, mod, mod0)
```

plotting 3 surrogate variables
```{r}
library(RColorBrewer)
colors <- brewer.pal(3, "Set2")
plot(svseq$sv, pch = 19, col=colors[batch])
text(svseq$sv, labels=sampNum, cex= 0.7)
```

Adding surrogate variables to actual design and testing using DESeq, using 3 surrogate variables
```{r}
ddssva <- dds
ddssva$SV1 <- svseq$sv[,1]
ddssva$SV2 <- svseq$sv[,2]
ddssva$SV3 <- svseq$sv[,3]
design(ddssva) <- ~ SV1 + SV2 + SV3 + batch
ddssva <- DESeq(ddssva)
```

Combat doesn't use the covariates from svseq, there are too many? Instead it uses the batch factor we gave earlier, 7 sibs, 5 11C.
Combat requires log counts, normalized or not?
The results are remarkably similar to sva output using limma removeBatchEffects function.
```{r ComBat}
# Loading all variables 
library(sva)
dataa <- log(counts(ddssva)+1)
modcombat = model.matrix(~1, data=sTable)
combatseq = ComBat(dat=dataa, batch=batch, mod = modcombat, par.prior = TRUE)
```

#Write combat output

# Writing csv with batch corrected counts
```{r}
combatseq = exp(combatseq) - 1
combatseq_counts <- ceiling(combatseq)
#write.csv(combatseq_counts, file = "/Volumes/cachannel/SCRIPTS/R/APRotations_RNA-Seq/stage18_APR_batchcorr/11C_Sibs/stage18_Sibs_11C_normalized_Counts_combat.csv")
```

# Combat Plot
Looks very similar to original plot, didn't remove any batches. This makes sense since it didn't use covariates from svseq, rather just the batch covariates. 
```{r combat plot}
project.pca <- prcomp(t((combatseq_counts)))
summary(project.pca)
project.pca.proportionvariances <- ((project.pca$sdev^2) / (sum(project.pca$sdev^2)))*100
```

```{r}
library(RColorBrewer)
colors <- brewer.pal(3, "Set2")

#Plots scatter plot for PC 1 and 2
plot(project.pca$x, type="n", main="Combat Plot", xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"), ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"))
points(project.pca$x, col=colors[batch], pch=19)
text(project.pca$x, labels=sampNum, cex= 0.7)
```


What does it look like for Sibs vs 11C (SVA)? 
There are now 2500 DE genes (it went increased?)
```{r}
res <- results(ddssva, alpha = 0.05, name = "batch_Sibs_vs_11C")
hist(res$pvalue, main = "batch_Sibs_vs_11C SVA PLOT", xlab = "p-values", col = "lavender")
summary(res)
#writeOutput(res, ddssva, "stage18", c("Sibs", "11C"), dir)
```

## LIMMA

```{r LIMMA}
suppressMessages(library(limma))

# get the sva covariates
sv_covar <- svseq$sv

counts_deseq_sva <- removeBatchEffect(log(counts(ddssva)+1), covariates = sv_covar)

project.pca <- prcomp(t(counts_deseq_sva))
summary(project.pca)
project.pca.proportionvariances <- ((project.pca$sdev^2) / (sum(project.pca$sdev^2)))*100
```

```{r}
#Plots scatter plot for PC 1 and 2
plot(project.pca$x, type="n", main="PCA Plot with SVA", xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"), ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"))
points(project.pca$x, col=colors[batch], pch=19)
text(project.pca$x, labels=sampNum, cex= 0.7)
```

# Writing csv with batch corrected counts
```{r write csv sva}
batch_corrected_counts_SVA <- exp(counts_deseq_sva) - 1
batch_corrected_counts_SVA <- ceiling(batch_corrected_counts_SVA)

#write.csv(batch_corrected_counts_SVA, file = "/Volumes/cachannel/SCRIPTS/R/APRotations_RNA-Seq/Batch_Correction/11C_Sibs/stage18_Sibs_11C_normalized_Counts_sva.csv")
```













