---
title: "PCA Plotting Function"
author: "LeAnn Lo"
date: "February 14, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PCA Plotting Function with VSD and RLD Correction

We use the PCA plotting in DESeq2 after transforming the data so the variance and mean relationship is linear.
Transform data so that we can visualize relationships between samples as well as perform clustering. Setting blind = FALSE as the dispersion estimations have already been done by running the DESeq() function. 
```{r}
#vsd_filtered <- vst(ddssva, blind=FALSE)
#rld_filtered <- rlog(ddssva, blind=FALSE)
```

Comapare the transformations. It appears the vsd transformation produces a more stable mena variance relationship.
```{r}
#library("vsn")
#meanSdPlot(assay(vsd_filtered))
#meanSdPlot(assay(rld_filtered))

```


PCA clustering function 
```{r}
# PCA plot of samples
PCAPlotter <- function(ntop, vsd) {
  # getting most variable genes 
  Pvars <- rowVars(assay(vsd))
  select <- order(Pvars, decreasing = TRUE)[seq_len(min(ntop, 
          length(Pvars)))]
  sampleNO <- sampNum
  # calculate pca - zero centering variables and scaling where all variables have unit variance
  PCA <- prcomp(t(assay(vsd)[select, ]), scale = T, center = TRUE)
  percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
  dataGG = data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2], 
                      PC3 = PCA$x[,3], PC4 = PCA$x[,4], 
                      condition = colData(vsd)$condition)
  # plotting
  print((qplot(PC1, PC2, data = dataGG, color =  condition, 
  	       main = paste("PC1 vs PC2, Top", toString(ntop), "Variable Genes"), size = I(6))
  	 + labs(x = paste0("PC1:", round(percentVar[1],4), "%"),
  	        y = paste0("PC2:", round(percentVar[2],4), "%"))
  	 + scale_colour_brewer(type="qual", palette=2) 
  	 + geom_label_repel(aes(label = sampleNO), color = "black")))
  PCA$rotation
  PCA$sdev
  return(PCA)
}
```

```{r}
PCA <- PCAPlotter(500, vsd_filtered)
PCAl <- PCAPlotter(1000, vsd_filtered)
PCA <- PCAPlotter(5000, vsd_filtered)
PCA <- PCAPlotter(10000, vsd_filtered)
PCA <- PCAPlotter(30000, vsd_filtered)
```

